---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R build status](https://github.com/trashbirdecology/parrottraits/workflows/R-CMD-check/badge.svg)](https://github.com/trashbirdecology/parrottraits/actions)
[![experimental](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## About \code{parrottraits}

The parrottraits R package provides the Parrot Trait Database, originally published in Burgio et al. (2018), as an internal data object, and provides minimal functionality for munging, visualizing, and interfacing with the data.

Planned functionality for \{code{parrottraits}} includes:  
- simple statistics and visualizations of both the traits database the phylogenetic supertree, as originally published in Burio et al. (2018)
- simple visualizations of trait distribution
- spatial distribution of traits over space
- integrating the taxonomic information used in the traits dataabase and supertree with other taxonomies

## Examples
Load traits database into the workspace:
```{r load-data}
data("traits")
dim(traits) # contains ~MANY~ traits  (columns!)
```



## Citations
__To cite the Parrot Traits database or companion R package (\code{parrottraits})__
```{r citation, echo=FALSE, eval=TRUE}
citation("parrottraits")[1] # data article
citation("parrottraits")[2] # R package 
```

## Installation

Install the development version from [GitHub](https://github.com/) with:

```{r install-packages, echo=TRUE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("trashbirdecology/parrottraits", force=FALSE)
```

## Code of Conduct
Please note that the parrottraits project is released with a [Contributor Code of Conduct](https://trashbirdecology.github.io/parrottraits//CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## USGS Disclaimers
*"These data are preliminary or provisional and are subject to revision. They are being provided to meet the need for timely best science. The data have not received final approval by the U.S. Geological Survey (USGS) and are provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the data."
"This information is preliminary or provisional and is subject to revision. It is being provided to meet the need for timely best science. The information has not received final approval by the U.S. Geological Survey (USGS) and is provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the information."
"This software is preliminary or provisional and is subject to revision. It is being provided to meet the need for timely best science. The software has not received final approval by the U.S. Geological Survey (USGS). No warranty, expressed or implied, is made by the USGS or the U.S. Government as to the functionality of the software and related material nor shall the fact of release constitute any such warranty. The software is provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the software."
*
