---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(parrottraits)
```

<!-- badges: start -->
![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)
<!-- badges: end --> 

## About 

The R package, \code{parrottraits}, provides quick access to the most recent version of the Parrot Traits Database, originally published in [Burgio et al. (2018)](https://www.sciencedirect.com/science/article/pii/S2352340919302331), and the original version. The Parrot Traits Database will be updated as new information becomes available. All are welcome to contribute new functional trait information, or suggest edits as necessary. 

In additon to the data, this package also provides minimal functionality for munging, visualizing, and interfacing with Parrot Trait Database.

__Planned functionality__

- simple statistics and visualizations of both the traits database the phylogenetic supertree, as originally published in Burio et al. (2018)
- simple visualizations of trait distribution 
- distribution of traits over space 
- integrating the taxonomic information used in the traits dataabase and supertree with other taxonomies

Load the most current version of the Parrot Traits database into  workspace:
```{r, echo=TRUE, eval=TRUE}
data("traits")
dim(traits)
```

```{r, include=FALSE}
data("traits_original")
n = length(setdiff(traits$taxon, traits_original$`Traits with inferred values`$species))
```

This package defaults to the most recent version of the Parrot Traits database ("traits"), which contains `r print(n)` species that are not available in the original data. 
```{r} 
# load the originallly published functional trait database
data("traits_original")
# see which species in the most recent version of "traits" were not originally published
sort(setdiff(traits$taxon, traits_original$`Traits with inferred values`$species))
```

To see what other data objects are available for import:
```{r, echo=TRUE, eval=FALSE}
data(package="parrottraits")
```



## Citations

**To cite this R package and explicitly updated versions of the Parrot Traits database, please use:**
```{r}
citation("parrottraits")[2]
```

**To cite the originally published Parrot Traits database and any subsequent versions, please use:**
```{r}
citation("parrottraits")[1]
```


## Installation

Install the development version from [GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
# devtools::install_github("trashbirdecology/parrottraits", force=FALSE)
```

## Code of Conduct

Please note that the parrottraits project is released with a [Contributor Code of Conduct](https://github.com/trashbirdecology/parrottraits/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.

## USGS Disclaimers
*“These data are preliminary or provisional and are subject to revision. They are being provided to meet the need for timely best science. The data have not received final approval by the U.S. Geological Survey (USGS) and are provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the data.” “This information is preliminary or provisional and is subject to revision. It is being provided to meet the need for timely best science. The information has not received final approval by the U.S. Geological Survey (USGS) and is provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the information.” “This software is preliminary or provisional and is subject to revision. It is being provided to meet the need for timely best science. The software has not received final approval by the U.S. Geological Survey (USGS). No warranty, expressed or implied, is made by the USGS or the U.S. Government as to the functionality of the software and related material nor shall the fact of release constitute any such warranty. The software is provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the software.” *
